Incremental λ-Calculus in Cache-Transfer Style

### 2017-07-11 popl18/source-language.tex
Use conditional compilation + add conditional remark

\Cref{fig:source-definition-base-terms-semantics} shows a
step-indexed big-step semantics for base terms, defined through
judgment~$\seval\sectx\sterm{n}\sclosedvalue$, pronounced ``Under
environment~$\sectx$, base term~$\sterm$ evaluates to closed
value~$\sclosedvalue$ in~$n$~steps.'' Our step-indexes count the number of
``nodes'' of a big-step derivation.\footnote{Instead,
  \citet{Ahmed2006stepindexed} and \citet{Acar08} count the
number of steps that small-step evaluation would take
% only for thesis:
(as we did in \cref{ch:bsos}),
%
but this simplifies some proof steps and makes a minor difference in others.}
We explain each rule in turn.
%
Rule \rulename{SVar} looks variable $\tx$ up in environment $\sectx$.
Other rules evaluate $\mathbf{let}$-binding $\slet{\ty = \ldots}{\sterm}$ in
environment $\sectx$: Each rule computes $\ty$'s new value $\sclosedvalue_\ty$ (taking~$m$ steps,
where $m$ can be zero) and evaluates in~$n$ steps body~$\sterm$ to
$\sclosedvalue$, using environment~$\sectx$ extended by binding~$\ty$
to~$\sclosedvalue_\ty$. The overall $\textbf{let}$-binding evaluates to
$\sclosedvalue$ in $m + n + 1$ steps. But different rules compute $\ty$'s value differently.
%
\rulename{STuple} looks each variable in~$\many\tx$ up in~$\sectx$ to evaluate
tuple~$\stuple{\many\tx}$ (in $m=0$ steps).
%
\rulename{SPrimitiveCall} evaluates function calls where
variable~$\tf$ is bound in~$\sectx$ to a primitive~$\sprim$.
How primitives are evaluated is specified by function $\tdelta\sprim\param$ from
closed values to closed values. To evaluate such a primitive call, this rule
applies~$\tdelta\sprim\param$ to~$\tx$'s value (in $m=0$ steps).
%
\rulename{SClosureCall} evaluates functions calls where variable~$\tf$ is bound
in~$\sectx$ to closure~$\sclosure{\sectx_\tf}{\slam{\tx}{\sterm_\tf}}$: this
rule evaluates closure body $\sterm_\tf$ in $m$
steps, using closure environment~$\sectx_\tf$ extended with the value
of argument~$\tx$ in~$\sectx$.

@@

If $\iectxlookup\iectx\tdf$ is a nil change. We have two rules
depending on the nature of $\iectxlookup\iectx\tf$. If
$\iectxlookup\iectx\tf$ is a closure, $\rulename{SDClosureNil}$
applies and in that case the nil change of~$\iectxlookup\iectx\tf$ is
the exact same closure. Hence, to compute $\tdy$, we must reevaluate
this closure applied to the updated
argument~$\iectxlookup{\newenv\iectx}\tx$ to a
value~$\sclosedvalue'_\ty$ and bind~$\tdy$ to
$\replace{\sclosedvalue'_\ty}$. In other words, this rule is
equivalent to $\rulename{SDReplaceCall}$ in the case where a closure
is replaced by itself.
%
If $\iectxlookup\iectx\tf$ is a primitive, $\rulename{SDPrimitiveNil}$
applies. The nil change of a primitive~$\sprim$ is its derivative
which interpretation is realized by a
function~$\Delta_\sprim(\param)$.  The evaluation of this function on
the input value and the input change leads to the
change~$\icloseddvalue_\ty$ bound to $\tdy$.

>>>>

\Cref{fig:source-definition-base-terms-semantics} shows a
step-indexed big-step semantics for base terms, defined through
judgment~$\seval\sectx\sterm{n}\sclosedvalue$, pronounced ``Under
environment~$\sectx$, base term~$\sterm$ evaluates to closed
value~$\sclosedvalue$ in~$n$~steps.'' Our step-indexes count the number of
``nodes'' of a big-step derivation.\footnote{Instead,
  \citet{Ahmed2006stepindexed} and \citet{Acar08} count the
number of steps that small-step evaluation would take%
\iftoggle{poplForThesis}{
(as we did in \cref{ch:bsos})%
}{}%
, but this simplifies some proof steps and makes a minor difference in others.}
We explain each rule in turn.
%
Rule \rulename{SVar} looks variable $\tx$ up in environment $\sectx$.
Other rules evaluate $\mathbf{let}$-binding $\slet{\ty = \ldots}{\sterm}$ in
environment $\sectx$: Each rule computes $\ty$'s new value $\sclosedvalue_\ty$ (taking~$m$ steps,
where $m$ can be zero) and evaluates in~$n$ steps body~$\sterm$ to
$\sclosedvalue$, using environment~$\sectx$ extended by binding~$\ty$
to~$\sclosedvalue_\ty$. The overall $\textbf{let}$-binding evaluates to
$\sclosedvalue$ in $m + n + 1$ steps. But different rules compute $\ty$'s value differently.
%
\rulename{STuple} looks each variable in~$\many\tx$ up in~$\sectx$ to evaluate
tuple~$\stuple{\many\tx}$ (in $m=0$ steps).
%
\rulename{SPrimitiveCall} evaluates function calls where
variable~$\tf$ is bound in~$\sectx$ to a primitive~$\sprim$.
How primitives are evaluated is specified by function $\tdelta\sprim\param$ from
closed values to closed values. To evaluate such a primitive call, this rule
applies~$\tdelta\sprim\param$ to~$\tx$'s value (in $m=0$ steps).
%
\rulename{SClosureCall} evaluates functions calls where variable~$\tf$ is bound
in~$\sectx$ to closure~$\sclosure{\sectx_\tf}{\slam{\tx}{\sterm_\tf}}$: this
rule evaluates closure body $\sterm_\tf$ in $m$
steps, using closure environment~$\sectx_\tf$ extended with the value
of argument~$\tx$ in~$\sectx$.

@@

If $\iectxlookup\iectx\tdf$ is a nil change. We have two rules
depending on the nature of $\iectxlookup\iectx\tf$. If
$\iectxlookup\iectx\tf$ is a closure, $\rulename{SDClosureNil}$
applies and in that case the nil change of~$\iectxlookup\iectx\tf$ is
the exact same closure. Hence, to compute $\tdy$, we must reevaluate
this closure applied to the updated
argument~$\iectxlookup{\newenv\iectx}\tx$ to a
value~$\sclosedvalue'_\ty$ and bind~$\tdy$ to
$\replace{\sclosedvalue'_\ty}$. In other words, this rule is
equivalent to $\rulename{SDReplaceCall}$ in the case where a closure
is replaced by itself.%
\begin{poplForThesis}
\footnote{Based on \cref{ch:bsos}, we are confident we
could in fact use the derivative of $\sterm_f$;}
\end{poplForThesis}
%
If $\iectxlookup\iectx\tf$ is a primitive, $\rulename{SDPrimitiveNil}$
applies. The nil change of a primitive~$\sprim$ is its derivative
which interpretation is realized by a
function~$\Delta_\sprim(\param)$.  The evaluation of this function on
the input value and the input change leads to the
change~$\icloseddvalue_\ty$ bound to $\tdy$.

### 2017-07-11 popl18/static-differentiation.tex
Use conditional compilation + add conditional remark

\subsection{Static differentiation in \source}
\label{sec:stat-diff-source}
Previous work~\citep{CaiEtAl2014ILC} defines differentiation for simply-typed
$\lambda$-calculus terms.
Figure~\ref{fig:popl-differentiation} shows differentiation for $\source$'s syntax.

>>>>

\subsection{Static differentiation in \source}
\label{sec:stat-diff-source}
\iftoggle{poplForThesis}{%
\Cref{def:derive} defines differentiation for simply-typed
$\lambda$-calculus terms.}{%
Previous work~\citep{CaiEtAl2014ILC} defines differentiation for simply-typed
$\lambda$-calculus terms.}
\Cref{fig:popl-differentiation} shows differentiation for $\source$ syntax.
